---
title: "Game Sales Markdown"
author: "Malik Hossain"
date: "2023-08-11"
output: html_document
---

## An Analysis on video game sales data across the big three console platforms
This analysis takes data from three different set from [this](https://www.kaggle.com/datasets/sidtwr/videogames-sales-dataset) Kaggle source and cleans each set for computing analyses on each and combining sets for clearer pictures

### Setting up the environment
Notes: Setting up my environment by loading the various packages used during the analysis and loading our datasets

```{r loading packages, message=FALSE, warning=FALSE}
library (tidyverse)
library(here)
library(ggplot2)
library(skimr)
library(janitor)
library(dplyr)
library(knitr)
```

```{r loading up datasets, echo=FALSE, paged.print=TRUE}
ps4_sales<-read.csv("PS4_GamesSales.csv")
print("PS4 Sales")
print(head(ps4_sales))

xbox_sales<-read.csv("XboxOne_GameSales.csv")
print("XBox Sales")
print(head(xbox_sales))

vg_sales<-read.csv("Video_Games_Sales_as_at_22_Dec_2016.csv")
print("General Video Game Sales")
print(head(vg_sales))
```

```{r Performing cleaning on the sets, echo=FALSE}
clean_ps4 <- ps4_sales %>% filter(Global != 0, Year != "N/A")
clean_xbox <- xbox_sales %>% filter(Global != 0, Year != "N/A")
clean_vg_sales <- vg_sales %>% filter(Global_Sales != 0, Year_of_Release != "N/A")
```

Visualizing PS4 sales by Year
```{r echo=FALSE}
yearly_ps4_sales<-clean_ps4 %>% group_by(Year)%>%summarize(sum(Global))
colnames(yearly_ps4_sales)[2] = "sales"

ggplot(data = yearly_ps4_sales) +
  geom_col(mapping = aes(x = Year, y = sales, fill = "sales"))
```

Visualizing XBox sales by Year
```{r echo=FALSE}
yearly_xbox_sales<-clean_xbox %>% group_by(Year)%>%summarize(sum(Global))
colnames(yearly_xbox_sales)[2] = "sales"

ggplot(data = yearly_xbox_sales) +
  geom_col(mapping = aes(x = Year, y = sales), fill = "cyan")
```
Comparing the two, we can see a clear lead for Ps4 sales during each year
```{r echo=FALSE}
sales_comp <- merge(x = yearly_ps4_sales, y = yearly_xbox_sales, by = "Year")
colnames(sales_comp)[2] = "Ps4_Sales"
colnames(sales_comp)[3] = "XBox_Sales"

ggplot(data = sales_comp) +
  geom_col(mapping = aes(x = Year, y = Ps4_Sales, fill = "ps4_Sales"))+
  geom_col(mapping = aes(x = Year, y = XBox_Sales, fill = "xBox_Sales") )+
  labs(y = "Sales", title = "Ps4 vs Xbox Game Sales")
```

Analyzing best selling genres of video games
```{r echo=FALSE}
genre_by_p <- clean_ps4 %>% group_by(Genre) %>% summarize(sum(Global)) 
colnames(genre_by_p)[2] = "sales"
print("PS4 sales by Genre")
genre_by_p %>% arrange(-sales)

genre_by_x <- clean_xbox %>% group_by(Genre) %>% summarize(sum(Global))
colnames(genre_by_x)[2] = "sales"
print("Xbox sales by Genre")
genre_by_x %>% arrange(-sales)
```
It seems Action games have more popularity on ps4 and shooters have most popularity on xbox although shooters are a very close second place for ps4








